---
title: "Affine MGARCH Simulation"
author: "Tianhao Li"
date: "2/8/2022"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Simulation function

**Notes**
**Inputs: theta = [lambda, omega, beta, alpha, gamma], rf, h_init, N, T**\
1. $\theta$:
  + $\lambda$: risk premium factor
  + $\omega, \beta, \alpha, \gamma$: GARCH parameters with respect to conditional variances.
2. rf: risk-free interest rate\
3. h_init: initial h value for generating h[i,j] matrix\
4. N: number of assets\
5. T: simulation time horizon\

**Outputs: simulated return, h and z**
1. simulated return: $R_t$
2. simulated z: random normal varaibles
3. h: 
```{r}
MHNGARCH_sim <- function(theta, rf, h_init, N, T){
    
    set.seed(123)
    
    # Generate z as a random normal matrix with N*T dimension
    z <- rnorm(N * T, 0, 1)
    z <- matrix(z, nrow = N)
    
    # Vectorize GARCH parameters
    w <- rep(theta$omega, N)
    beta <- rep(theta$beta, N)
    alpha <- rep(theta$alpha, N)
    gamma <- rep(theta$gamma, N)
    
    # Generate lambda matrix as N*N dimension
    lambda <- rep(theta$lambda, N*N)
    lambda.matrix <- matrix(lambda, nrow = N)
    
    # Determine A_NN matrix
    A <- rep(theta$a, N*N)
    A.matrix <- matrix(A, nrow = N)
    
    # Generate the first column of h
    h <- matrix(rep(0, N*T), nrow = N)
    h[, 1] <- rep(h_init, N)
    
    # Get h matrix as GARCH time series with all parameters
    for (t in 2:T) {
    h[, t] <- w + beta * h[, t-1] + alpha * (z[, t-1] - gamma * sqrt(h[, t-1]))^2
    }
    
    # Generate epsilon as a random normal matrix
    epsilon <- sqrt(h) * z
    
    # Get simulated return
    ret <- rf + lambda.matrix %*% h + A.matrix %*% epsilon
    ret_sim <- matplot(t(ret), type = "l")
    
  return(ret_sim)


}
```

```{r test}
# Testify the simulation function
theta_test <- list(lambda = 4, omega = 8e-5, alpha = 6e-5,
             beta = 0.7, gamma = 0, a = 1)

MHNGARCH_sim(theta_test, rf = 0, h_init = 1e-04, N = 4, T = 5000)


```